Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/advup.tex

The discretized $Q$-function rewrites as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way of approximating $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or $Q$-learning to a reparameterized $Q$-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta = (\theta, \pi)$. To avoid cumbersome notations, in what follows,
parameter indices are dropped.  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independantly of $\deltat$, this parameterization
reasonably scales the contribution of actions relative to states in the value
of $Q$.  Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
However, the proposed reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$\footnote{
	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is the same.
}.
Indeed, for any given pair $V$, $A$,
\begin{align}
	\tilde{V}(s) &= V(s) - f(s)\\
	\tilde{A}(s, a) &= A(s, a) + \frac{f(s)}{\deltat}
\end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A = A^\pi_\deltat$ and $V = V^\pi_\deltat$.
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) = \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper ranking of actions can be derived for any $\deltat$.

>>>>

The discretized $Q$-function rewrites as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way to approximate $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or $Q$-learning to a reparameterized $Q$-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta = (\theta, \pi)$. To avoid cumbersome notation, in what follows,
parameter indices are dropped.  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independantly of $\deltat$, this parameterization
provides reasonable scaling of the contribution of actions versus states
in $Q$.
Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
Still, this reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$\footnote{
	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is
	the same. \TODO{still unclear why we'd care about identifiability}
}.
Indeed, for any given pair $(V,\,A)$, the pair $(V(s) -
f(s),\,A(s,a)+f(s)/\deltat)$ (for an arbitrary $f$)
% \begin{align}
% 	\tilde{V}(s) &= V(s) - f(s),\qquad
% 	\tilde{A}(s, a) = A(s, a) + \frac{f(s)}{\deltat}
% \end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A =
A^\pi_\deltat$ and $V = V^\pi_\deltat$: indeed
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) \deq \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper ranking of actions can be derived for any $\deltat$.
